@page

@section  Association Class
@pdfsection{Association Class}
The @code{Association} class is an abstract class which serves the
sole purpose of grouping several subclasses which have common
functionality.  There is no specific class or instance methods
directly implemented by this class, however, some methods are documented
here because of their common interface with all subclasses of the
@code{Association} class.

The subclasses of this class (@code{LookupKey},
@code{ObjectAssociation}, @code{StringAssociation} and
@iftex @break @end iftex
@code{IntegerAssociation}) are used to associate a key with a value and
provide a means for manipulating these associations.  Instances of these
classes are principally used by the @code{Set} and @code{Dictionary}
classes but may also be useful wherever a key/value pair may be useful.


@subsection Association Class Methods
@pdfsubsection{Association Class Methods}
There are no class methods implemented by this class.


@subsection Association Instance Methods
@pdfsubsection{Association Instance Methods}
There are no class methods implemented by this class, however, some
methods are documented here because of their common interface with all
subclasses of the @code{Association} class.  These methods are actually
implemented by all subclasses of this class.






@pdfsubsubsection {Compare}
@deffn {Compare} Compare::Association
@sp 2
@example
@group
r = gCompare(i, obj);

object  i;
object  obj;
int     r;
@end group
@end example
This method is used by the generic container classes to determine
the equality of the keys represented by @code{i} and @code{obj}. 
@code{r} is -1 if the value represented by @code{i} is less than
the value represented by @code{obj}, 1 if the value of @code{i}
is greater than @code{obj}, and 0 if they are equal.
@c @example
@c @group
@c @exdent Example:
@c
@c @end group
@c @end example
@sp 1
See also:  @code{Hash::Association}
@end deffn










@pdfsubsubsection {Hash}
@deffn {Hash} Hash::Association
@sp 2
@example
@group
val = gHash(i);

object  i;
int     val;
@end group
@end example
This method is used by the generic container classes to obtain hash
values for the key.  @code{val} is a hash value between 0 and a large
integer value.
@c @example
@c @group
@c @exdent Example:
@c
@c @end group
@c @end example
@sp 1
See also:  @code{Compare::Association}
@end deffn












