@page

@section LinkObjectSequence Class
This class is a subclass of @code{Sequence} and is used to provide a
mechanism to enumerate through all the objects on the links in a linked
list structure without effecting the structure.

The linked list structure is usually represented by instances of the
@code{LinkObject} class.  Typically, the @code{Sequence} instance method
of this class is used to create the instances of the
@iftex 
@break 
@end iftex
@code{LinkObjectSequence} class.



@subsection LinkObjectSequence Class Methods
There is only one class method associated with this class.  It is used
to create new instances of itself and is normally only called by
sequence methods associated with the @code{LinkObject} class.





@deffn {NewWithObj} NewWithObj::LinkObjectSequence
@sp 2
@example
@group
i = gNewWithObj(LinkObjectSequence, lnk);

object  lnk;
object  i;
@end group
@end example
This class method creates instances of the @code{LinkObjectSequence}
class.  It is normally only called by the sequence method associated
with the @code{LinkObject} class.  @code{lnk} is the first link in the
list, and @code{i} is the sequence instance.
@c @example
@c @group
@c @exdent Example:
@c 
@c @end group
@c @end example
@sp 1
See also:  @code{Next::LinkObjectSequence, Sequence::LinkObject}
@end deffn




@subsection LinkObjectSequence Instance Methods
This class only has a single instance method which is used to enumerate
through object on a list.






@deffn {Next} Next::LinkObjectSequence
@sp 2
@example
@group
obj = gNext(i);

object  i;
object  obj;
@end group
@end example
This method is used to enumerate through the objects contained within
the links on a linked list.  Each time @code{Next} is called the
following object on the linked list is returned.  It does this in a
non-destructive way so the associated linked list is not effected.

When @code{Next} is called after the last object has been returned (or
if there are no links on the list) it will return @code{NULL} and
automatically dispose of the sequence object @code{i}.  Therefore,
the only time @code{Dispose} would be needed (if the garbage collector
weren't being used) would be if the entire list was not enumerated.
@example
@group
@exdent Example:

object  ll;  /*  linked object list */
object  s;   /*  link sequence      */
object  obj; /*  an object          */

/*  ll must be initialized previously  */

for (s=gSequence(ll) ; obj = gNext(s) ; )  @{
        /*  do something with object obj  */
@}
@end group
@end example
@sp 1
See also:  @code{Sequence::LinkObject}
@end deffn






