@page

@section LinkSequence Class
This class is a subclass of @code{Sequence} and is used to provide a
mechanism to enumerate through all the links in a linked list structure
without effecting the structure.

The linked list structure is usually represented by instances of the
@code{LinkList} or @code{LinkObject} classes.  Typically, instance
methods of these classes are used to create the instances of the
@code{LinkSequence} class.  These methods have names which begin with
@code{Sequence}.



@subsection LinkSequence Class Methods
There is only one class method associated with this class.  It is used
to create new instances of itself and is normally only called by
sequence methods associated with the @code{LinkList} or @code{LinkObject}
classes.





@deffn {NewWithObj} NewWithObj::LinkSequence
@sp 2
@example
@group
i = gNewWithObj(LinkSequence, lnk);

object  lnk;
object  i;
@end group
@end example
This class method creates instances of the @code{LinkSequence} class.
It is normally only called by sequence methods associated with the
@code{LinkList} or @code{LinkObject} classes.  @code{lnk} is the first
link in the list, and @code{i} is the sequence instance.
@c @example
@c @group
@c @exdent Example:
@c 
@c @end group
@c @end example
@sp 1
See also:  @code{Next::LinkSequence, Sequence::LinkList}
@end deffn




@subsection LinkSequence Instance Methods
This class only has a single instance method which is used to enumerate
through links on a list.






@deffn {Next} Next::LinkSequence
@sp 2
@example
@group
lnk = gNext(i);

object  i;
object  lnk;
@end group
@end example
This method is used to enumerate through the links on a linked list.
Each time @code{Next} is called the following link on the linked list
is returned.  It does this in a non-destructive way so the associated
linked list is not effected.

When @code{Next} is called after the last link has been returned (or
if there are no links on the list) it will return @code{NULL} and
automatically dispose of the sequence object @code{i}.  Therefore,
the only time @code{Dispose} would be needed (if the garbage collector
weren't being used) would be if the entire list was not enumerated.
@example
@group
@exdent Example:

object  ll;  /*  linked list    */
object  s;   /*  link sequence  */
object  lnk; /*  one link       */

/*  ll must be initialized previously  */

for (s=gSequence(ll) ; lnk = gNext(s) ; )  @{
        /*  do something with link lnk  */
@}
@end group
@end example
@sp 1
See also:  @code{Sequence::LinkList, SequenceLinks::LinkObject}
@end deffn











