@page

@section LongInteger Class
The @code{LongInteger} class is used to represent the C language
@code{long} data type as a Dynace object.  It is a subclass of the
@code{Number} class.  Even though the @code{LongInteger} class
implements most of its own functionality, it is documented as part of
the @code{Number} class because most of the interface is the same for
all subclasses of the @code{Number} class.  Differences are documented
in this section.





@subsection LongInteger Class Methods
The @code{LongInteger} class has only one class method and it is used to
create new instances of itself.



@deffn {NewWithLong} NewWithLong::LongInteger
@sp 2
@example
@group
i = gNewWithLong(LongInteger, val);

long    val;
object  i;
@end group
@end example
This class method creates instances of the LongInteger class.  @code{val}
is the initial value of the integer being represented.  

The value returned is a Dynace instance object which represents the integer
passed.

Note that the default disposal methods are used by this class since
there are no special storage allocation requirements.
@example
@group
@exdent Example:

object  x;

x = gNewWithLong(LongInteger, 55L);
@end group
@end example
@sp 1
See also:  @code{LongValue::Number, Dispose::Object}
@end deffn





@subsection LongInteger Instance Methods
The instance methods associated with the @code{LongInteger} class
provide a means of changing and obtaining the value associated with an
instance of the @code{LongInteger} class.  All of the
@code{LongInteger} class instance methods are documented in the
@code{Number} class because of their common interface with all other
subclasses of the @code{Number} class.

















