@page

@section SetSequence Class
@pdfsection{SetSequence Class}
This class is a subclass of @code{Sequence} and is used to provide a
mechanism to enumerate through all the objects in a @code{Set} (or one
of its subclasses) without effecting the structure.

Typically, the @code{Sequence} instance method of the @code{Set} class
is used to create the instances of the @code{SetSequence} class.  



@subsection SetSequence Class Methods
@pdfsubsection{SetSequence Class Methods}
There is only one class method associated with this class.  It is used
to create new instances of itself and is only called by the @code{Sequence}
method associated with the @code{Set} class.

@subsection SetSequence Instance Methods
@pdfsubsection{SetSequence Instance Methods}
This class only has a single instance method which is used to enumerate
through objects in the @code{Set}.






@pdfsubsubsection {Next}
@deffn {Next} Next::SetSequence
@sp 2
@example
@group
obj = gNext(i);

object  i;
object  obj;
@end group
@end example
This method is used to enumerate through all the objects in a @code{Set}.
Each time @code{Next} is called the following object in the @code{Set}
is returned.  It does this in a non-destructive way so the associated
@code{Set} is not effected.

When @code{Next} is called after the last object has been returned (or
if there are no objects in the @code{Set}) it will return @code{NULL} and
automatically dispose of the sequence object @code{i}.  Therefore,
the only time @code{Dispose} would be needed (if the garbage collector
weren't being used) would be if the entire list was not enumerated.

Note that if this method is used with one of the @code{Dictionary} classes
each call to @code{Next} returns the appropriate @code{Association} type.
The @code{gValue} and @code{gKey} generics may be used to obtain
the desired values from the @code{Association} returned.
@example
@group
@exdent Example:

object  set;  /*  Set                        */
object  s;    /*  Set sequence               */
object  obj;  /*  one object or association  */

/*  set must be initialized previously  */

for (s=gSequence(set) ; obj = gNext(s) ; )  @{
        /*  do something with object obj  */
@}
@end group
@end example
@sp 1
See also:  @code{Sequence::Set}
@end deffn



