@page

@section ShortInteger Class
@pdfsection{ShortInteger Class}
The @code{ShortInteger} class is used to represent the C language
@code{short} data type as a Dynace object.  It is a subclass of the
@code{Number} class.  Even though the @code{ShortInteger} class
implements most of its own functionality, it is documented as part of
the @code{Number} class because most of the interface is the same for
all subclasses of the @code{Number} class.  Differences are documented
in this section.


@subsection ShortInteger Class Methods
@pdfsubsection{ShortInteger Class Methods}
The @code{ShortInteger} class has only one class method and it is used
to create new instances of itself.



@pdfsubsubsection {NewWithInt}
@deffn {NewWithInt} NewWithInt::ShortInteger
@sp 2
@example
@group
i = gNewWithInt(ShortInteger, s);

int     s;
object  i;
@end group
@end example
This class method creates instances of the ShortInteger class.  @code{s}
is the initial value of the integer being represented.  Note that
@code{s} is of type @code{int}.  That is because if you pass a short
type to a generic function (which uses a variable argument declaration)
the C language will automatically promote it to an @code{int}.

The value returned is a Dynace instance object which represents the integer
passed.

Note that the default disposal methods are used by this class since
there are no special storage allocation requirements.
@example
@group
@exdent Example:

object  x;

x = gNewWithInt(ShortInteger, 55);
@end group
@end example
@sp 1
See also:  @code{ShortValue::Number, Dispose::Object}
@end deffn





@subsection ShortInteger Instance Methods
@pdfsubsection{ShortInteger Instance Methods}
The instance methods associated with the @code{ShortInteger} class
provide a means of changing and obtaining the value associated with an
instance of the @code{ShortInteger} class.  All of the
@code{ShortInteger} class instance methods are documented in the
@code{Number} class because of their common interface with all other
subclasses of the @code{Number} class.















