@page

@section UnsignedShortInteger Class
The @code{UnsignedShortInteger} class is used to represent the C language
@code{unsigned short} data type as a Dynace object.  It is a subclass of the
@code{Number} class.  Even though the @code{UnsignedShortInteger} class
implements most of its own functionality, it is documented as part of
the @code{Number} class because most of the interface is the same for
all subclasses of the @code{Number} class.  Differences are documented
in this section.


@subsection UnsignedShortInteger Class Methods
The @code{UnsignedShortInteger} class has only one class method and it
is used to create new instances of itself.



@deffn {NewWithUnsigned} NewWithUnsigned::UnsignedShortInteger
@sp 2
@example
@group
i = gNewWithUnsigned(UnsignedShortInteger, s);

unsigned s;
object  i;
@end group
@end example
This class method creates instances of the @code{UnsignedShortInteger}
class.  @code{s} is the initial value of the integer being represented.
Note that @code{s} is of type @code{unsigned}.  That is because if you
pass an unsigned short type to a generic function (which uses a variable
argument declaration) the C language will automatically promote it to an
@code{unsigned}.

The value returned is a Dynace instance object which represents the
integer passed.

Note that the default disposal methods are used by this class since
there are no special storage allocation requirements.
@example
@group
@exdent Example:

object  x;

x = gNewWithUnsigned(UnsignedShortInteger, 55);
@end group
@end example
@sp 1
See also:  @code{UnsignedShortValue::Number, Dispose::Object}
@end deffn





@subsection UnsignedShortInteger Instance Methods
The instance methods associated with the @code{UnsignedShortInteger}
class provide a means of changing and obtaining the value associated
with an instance of the 
@iftex
@hfil @break 
@end iftex
@code{UnsignedShortInteger} class.
All of the @code{UnsignedShortInteger} class instance methods are
documented in the @code{Number} class because of their common interface
with all other subclasses of the @code{Number} class.

















